---
title: "Cross-Correlation Analyses (Figure S6)"   
author:  | 
  | Nicole Nova, Ethan R. Deyle, Marta S. Shocket, Andrew J. MacDonald, Marissa L. Childs, Martin Rypdal, George Sugihara, and Erin A. Mordecai
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

## Cross-Correlation Analyses (Figure S6)

If necessary, we can clear the environment to start with a clean workspace.
```{r}
rm(list=ls())
```

Help functions.
```{r}
source("data_formatting.R")
```

Here, we can load previous results.
```{r}
#load(file="../cache/Figure_S6.RData") 
```


### Data

We load and process the data and create our main dataframe PR, which stands for "Puerto Rico."
```{r}
PR <- data.clean("combined_SanJuan_PuertoRico_1990-2009")
```

Add lagged and averaged lagged variables.
```{r}
PR <- data.process(PR)
```

Only use the four variables.
```{r}
PR <- PR[,3:6]
```


### Cross-Correlation and Figure S6

```{r}
plot.xcor <- function(driver, mainTitle){
  
  par(mar = c(3.5,3.5,1.7,1), mgp = c(2,0.5,0), oma = c(0,0,0,0))
  
  # Cross-correlation
  xcorr_cases_driver <- ccf(PR$cases, na.omit(PR[[driver]]), lag.max = NULL, 
                            type = c("correlation", "covariance"),
                            plot = TRUE, main = "", axes = FALSE, xlab = "", ylab = "")
  box(col="black")
  axis(1, col="black", col.ticks="black", col.axis="black", cex.axis=1.2, tck = 0.02)
  mtext("Time lag (week)", side=1, line=2.2, col="black", cex=1.4)
  
  axis(2, col="black", col.ticks="black", col.axis="black", cex.axis=1.2, tck = 0.02)
  mtext(expression(paste("Cross-correlation with incidence (",rho,")")), side=2, line=2, 
        col="black", cex=1.4)
  
  bio_lag <- xcorr_cases_driver$lag[which(xcorr_cases_driver$acf == max(xcorr_cases_driver$acf))]
  abline(v = bio_lag, col = "red", lwd = 2)
  title(mainTitle, cex.main=1.4)
  
  file_name = paste("../output/fig_S6/", names(PR[driver]), ".pdf", sep="")
  dev.copy(pdf, file=file_name, width = 5, height = 5)
  dev.off()
}

# Temp and Incidence
plot.xcor("temp", expression(paste("Temperature")))

# Rainfall and Incidence
plot.xcor("prec", expression(paste("Rainfall")))

# Susceptibles Index and Incidence
plot.xcor("mu", expression(paste("Susceptibles Index (", lambda,")")))
```

Here, we can save our progress.
```{r}
save.image(file="../cache/Figure_S6.RData") 
```