---
title: "Evidence of Nonlinear Dynamics (Figure S5)"   
author:  | 
  | Nicole Nova, Ethan R. Deyle, Marta S. Shocket, Andrew J. MacDonald, Marissa L. Childs, Martin Rypdal, George Sugihara, and Erin A. Mordecai
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

## Evidence of Nonlinear Dynamics (Figure S5)

If necessary, we can clear the environment to start with a clean workspace.
```{r}
rm(list=ls())
```

Help functions.
```{r}
source("data_formatting.R")
```

Here, we can load previous results.
```{r}
#load(file="../cache/Figure_S5.RData") 
```


### Data

We load and process the data and create our main dataframe PR, which stands for "Puerto Rico."
```{r}
PR <- data.clean("combined_SanJuan_PuertoRico_1990-2009")
```

Add lagged and averaged lagged variables.
```{r}
PR <- data.process(PR)
```

Normalize the variables before performing EDM analyses.
```{r}
# Normalize the whole dataset
PR.norm <- scale(PR[,c(-1,-2)], center = TRUE, scale = TRUE)

# Convert back to dataframe
PR.norm <- as.data.frame(PR.norm)

# Add the date back
PR.norm <- cbind(PR[1], PR[2], PR.norm)

# Renaming PR to PR.norm since that's what we'll be using in the next EDM analyses
PR.orig <- PR
PR <- PR.norm
```

Only use the four variables.
```{r}
PR <- PR[,3:6]
```


### S-map

Here we are distinguishing nonlinear dynamical systems from linear stochastic systems and quantifying the nonlinearity. We will use the S-map method to find the value of theta for each variable. We will use the optimal embedding dimensions found from the simplex projection analysis in the previous section.
```{r}
em <- c(8,8,6,5) # Optimal embedding dimensions from previous analyses

E_cases <- 8
E_temp <- 8
E_prec <- 6
E_mu <- 5
```

```{r}
# S-map
smap.list <- list()
for (i in 1:length(PR)) {
  (sm <- s_map(PR[[i]], E = get(paste("E_", names(PR)[i], sep="")), 
               tau = 6, exclusion_radius = 6, tp = 6))
  smap.list[[i]] <- sm
}
```


### Figure S5

```{r}
par(mar = c(3.3,6,1.4,3.5), mgp = c(2,0.5,0), oma = c(0,0,0,0))

lb <- c("Incidence", "Temperature", "Rainfall", expression(paste("Susceptibles Index (", lambda,")")))

# Plot predictive skill (rho) vs state-dependency parameter (theta)
for (i in 1:length(PR)) {
  plot(smap.list[[i]]$theta, smap.list[[i]]$rho, type="l", lwd=3, col="red",
       axes = FALSE,
       xlab = "", ylab = "")
  box(col="black")
  axis(1, col="black", col.ticks="black", col.axis="black", cex.axis=1, tck = 0.02)
  mtext(expression(paste("Nonlinearity (",theta,")")), side=1, line=2, col="black", cex=1.2)
  
  axis(2, col="black", col.ticks="black", col.axis="black", cex.axis=1, tck = 0.02)
  mtext(expression(paste("Correlation coefficient (",rho,")")), side=2, line=2, col="black", cex=1.2)

  abline(v = smap.list[[i]]$theta[which.max(smap.list[[i]]$rho)], col="black", lwd=3, lty=2)
  title(lb[i])
  
  file_name = paste("../output/fig_S5/smap_", names(PR[i]), ".pdf", sep="")
  dev.copy(pdf, file=file_name, width = 5, height = 4)
  dev.off()
}
```

Here, we can save our progress.
```{r}
save.image(file="../cache/Figure_S5.RData") 
```