---
title: "Time Series Data (Figure 1)"   
author:  | 
  | Nicole Nova, Ethan R. Deyle, Marta S. Shocket, Andrew J. MacDonald, Marissa L. Childs, Martin Rypdal, George Sugihara, and Erin A. Mordecai
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

## Time Series Data (Figure 1)

If necessary, we can clear the environment to start with a clean workspace.
```{r}
rm(list=ls())
```

Help functions.
```{r}
source("data_formatting.R")
```

Here, we can load previous results.
```{r}
#load(file="../cache/Figure_1.RData") 
```


### Data

We load and process the data and create our main dataframe PR, which stands for "Puerto Rico."
```{r}
PR <- data.clean("combined_SanJuan_PuertoRico_1990-2009")
```


### Figure 1

Next, we will add a date variable that is formatted for whole time series analyses.
```{r}
PR$Year <- as.Date(as.character(levels(PR$week_start_date)), 
                     format = "%m/%d/%Y")[PR$week_start_date]
```
 
Clean up order of columns.
```{r}
# Create an array containing all the variable names 
vbs <- c("cases", "temp", "prec", "mu")

PR <- PR[, c("Year", "week", vbs)]
```

We visualize the time series data. 
```{r echo=T, messages=F, warning=F, results='hide',fig.keep='all', fig.width=8,fig.height=8}

cases_ts <- ggplot(PR, aes(Year, cases)) +
  geom_line() + 
  ylim(0,500) +
  ylab("Incidence") +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x=element_blank())

temp_ts <- ggplot(PR, aes(Year, temp)) +
  geom_line(color='red') +
  ylim(22,30) +
  ylab("Temperature (°C)") +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x=element_blank())

prec_ts <- ggplot(PR, aes(Year, prec)) +
  geom_line(color='blue') +
  ylim(0,300) +
  ylab("Rainfall (mm)") +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x=element_blank())
  
mu_ts <- ggplot(PR, aes(Year, mu)) +
  geom_line(color='purple') +
  ylim(0.5,1.5) +
  ylab(expression(paste("Susceptibles Index (", lambda,")"))) +
  xlab("Year") +
  theme(axis.title.x = element_text(), axis.text.x = element_text()) 
  
grid.newpage()
g_cases <- ggplotGrob(cases_ts)
g_temp <- ggplotGrob(temp_ts)
g_prec <- ggplotGrob(prec_ts)
g_mu <- ggplotGrob(mu_ts)
g2 <- rbind(g_cases, g_temp, g_prec, g_mu, size = "last")
grid.draw(g2)

ggsave("../output/fig_1/Figure_1.pdf", g2, width = 8, height = 8)
```

Here, we can save our progress.
```{r}
save.image(file="../cache/Figure_1.RData") 
```