---
title: "Dengue Forecasting Challenge: Testing (Table S1 and Figures S9 and S11)"   
author:  | 
  | Nicole Nova, Ethan R. Deyle, Marta S. Shocket, Andrew J. MacDonald, Marissa L. Childs, Martin Rypdal, George Sugihara, and Erin A. Mordecai
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

## Dengue Forecasting Challenge: Testing (Table S1 and Figures S9 and S11)

If necessary, we can clear the environment to start with a clean workspace.
```{r}
rm(list=ls())
```

Help functions and packages.
```{r}
library("fGarch")
library("fitdistrplus")

source("data_formatting.R")
source("forecast_challenge_functions.R")
```

Here, we can load previous results.
```{r}
#load(file="../cache/Table_S1_and_Figures_S9_S11.RData") 
```


### State-Space Reconstructed Forecasting Model

```{r}
# The best SSR model embedding
ts.dim.cases.drivers <- c("cases", "cases.lag_1", "cases.lag_2", "cases.lag_3", "cases.lag_4",
                          "mu.lag_5", "temp.lag_9", "prec.avglag_3")
```


### Testing Data

We load and process the data and create our main dataframe PR, which stands for "Puerto Rico."
```{r}
PR <- data.clean("combined_SanJuan_PuertoRico_1990-2013")
```

Add lagged and averaged lagged variables.
```{r}
PR <- data.process(PR)
```

Normalize the variables before performing EDM analyses.
```{r}
# Normalize the whole dataset
PR.norm <- scale(PR[,c(-1,-2)], center = TRUE, scale = TRUE)

# Convert back to dataframe
PR.norm <- as.data.frame(PR.norm)

# Add the date back
PR.norm <- cbind(PR[1], PR[2], PR.norm)

# Renaming PR to PR.norm since that's what we'll be using in the next EDM analyses
PR.orig <- PR
PR <- PR.norm
```

Forecasting targets:
1. Peak week (bin by week)
2. Peak incidence (bin by 50 and over 500)
3. Total cases (bin by 1000 and over 10,000)

First we obtain the observed cases for each testing season (2009/2010-2012/2013).
```{r}
obs.s <- PR$cases*sd(PR.orig$cases) + mean(PR.orig$cases)
start.season1 <- 52*19 + 1
start.season2 <- 52*20 + 1
start.season3 <- 52*21 + 1
start.season4 <- 52*22 + 1
obs.s1 <- obs.s[start.season1:(start.season2-1)]
obs.s2 <- obs.s[start.season2:(start.season3-1)]
obs.s3 <- obs.s[start.season3:(start.season4-1)]
obs.s4 <- obs.s[start.season4:length(obs.s)]
```


### Table S1 (2009/2010-2012/2013)


#### Peak week

Week of max incidence obtained for each 1-7 forecast:
```{r}
forecast.seasons.peak.wk <- forecast.test(PR, 988, ts.dim.cases.drivers, 1)
```

```{r}
max.wk.s1 <- which.max(obs.s1) # 43
max.wk.s2 <- which.max(obs.s2) # 16
max.wk.s3 <- which.max(obs.s3) # 20
max.wk.s4 <- which.max(obs.s4) # 32
```

Now we will calculate the scores.
```{r}
wk.score.1 <- calc.score.peak.wk(forecast.seasons.peak.wk, 1, obs.s1)
# -5.035549 
wk.score.2 <- calc.score.peak.wk(forecast.seasons.peak.wk, 2, obs.s2)
# -1.597884
wk.score.3 <- calc.score.peak.wk(forecast.seasons.peak.wk, 3, obs.s3)
# -2.607911
wk.score.4 <- calc.score.peak.wk(forecast.seasons.peak.wk, 4, obs.s4) 
# -3.180466
```

Total score for peak weak:
```{r}
score.peak.wk <- mean(c(wk.score.1, wk.score.2, wk.score.3, wk.score.4))
score.peak.wk
# -3.105452
``` 
Table S1: -3.11


#### Peak incidence

Max incidence obtained for each 1-7 forecast:
```{r}
forecast.seasons.peak.inc <- forecast.test(PR, 988, ts.dim.cases.drivers, 2)
```

```{r}
max.inc.s1 <- max(obs.s1) # Bin 50-100
max.inc.s2 <- max(obs.s2) # Bin 250-300
max.inc.s3 <- max(obs.s3) # Bin 50-100
max.inc.s4 <- max(obs.s4) # Bin 200-250
```

```{r}
inc.score.1 <- calc.score.peak.tot.inc(forecast.seasons.peak.inc, 1, max(obs.s1), 50, 100, 500)
# -0.8737035
inc.score.2 <- calc.score.peak.tot.inc(forecast.seasons.peak.inc, 2, max(obs.s2), 250, 300, 500)
# -0.7797399
inc.score.3 <- calc.score.peak.tot.inc(forecast.seasons.peak.inc, 3, max(obs.s3), 50, 100, 500)
# -0.6566172
inc.score.4 <- calc.score.peak.tot.inc(forecast.seasons.peak.inc, 4, max(obs.s4), 200, 250, 500)
# -1.756885
```

Total score for peak incidence:
```{r} 
peak.inc.score <- mean(c(inc.score.1, inc.score.2, inc.score.3, inc.score.4)) 
peak.inc.score
# -1.016736
``` 
Table S1: -1.02


#### Seasonal incidence

Total incidence obtained for each 1-7 forecast:
```{r}
forecast.seasons.tot.inc <- forecast.test(PR, 988, ts.dim.cases.drivers, 3)
```

```{r}
tot.inc.s1 <- sum(obs.s1) # Bin 2000-3000
tot.inc.s2 <- sum(obs.s2) # Bin 4000-5000
tot.inc.s3 <- sum(obs.s3) # Bin 1000-2000
tot.inc.s4 <- sum(obs.s4) # Bin 5000-6000
```

```{r}
tot.inc.score.1 <- calc.score.peak.tot.inc(forecast.seasons.tot.inc, 1, sum(obs.s1), 2000, 3000, 10000) 
# -1.494832
tot.inc.score.2 <- calc.score.peak.tot.inc(forecast.seasons.tot.inc, 2, sum(obs.s2), 4000, 5000, 10000) 
# -1.175615
tot.inc.score.3 <- calc.score.peak.tot.inc(forecast.seasons.tot.inc, 3, sum(obs.s3), 1000, 2000, 10000) 
# -0.3744226
tot.inc.score.4 <- calc.score.peak.tot.inc(forecast.seasons.tot.inc, 4, sum(obs.s4), 5000, 6000, 10000) 
# -1.997724
```

```{r}
tot.inc.score <- mean(c(tot.inc.score.1, tot.inc.score.2, tot.inc.score.3, tot.inc.score.4))
tot.inc.score
# -1.260648
```
Table S1: -0.26


Table S1:
```{r}
peak.inc.score
score.peak.wk
tot.inc.score
```


### Figure S9
 
```{r}
#boxplot.target(forecast.seasons, seasonNumber, obs.sx, fig, target)

# Peak week
boxplot.target(forecast.seasons.peak.wk, 1, obs.s1, "fig_S9", 1)
boxplot.target(forecast.seasons.peak.wk, 2, obs.s2, "fig_S9", 1)
boxplot.target(forecast.seasons.peak.wk, 3, obs.s3, "fig_S9", 1)
boxplot.target(forecast.seasons.peak.wk, 4, obs.s4, "fig_S9", 1)

# Peak incidence
boxplot.target(forecast.seasons.peak.inc, 1, obs.s1, "fig_S9", 2)
boxplot.target(forecast.seasons.peak.inc, 2, obs.s2, "fig_S9", 2)
boxplot.target(forecast.seasons.peak.inc, 3, obs.s3, "fig_S9", 2)
boxplot.target(forecast.seasons.peak.inc, 4, obs.s4, "fig_S9", 2)

# Seasonal incidence
boxplot.target(forecast.seasons.tot.inc, 1, obs.s1, "fig_S9", 3)
boxplot.target(forecast.seasons.tot.inc, 2, obs.s2, "fig_S9", 3)
boxplot.target(forecast.seasons.tot.inc, 3, obs.s3, "fig_S9", 3)
boxplot.target(forecast.seasons.tot.inc, 4, obs.s4, "fig_S9", 3)
```


### Figure S11

Get the forecasts.
```{r}
forecast.seasons <- forecast.test(PR, 988, ts.dim.cases.drivers, 0)
```

Plot the forecasts for all seasons.
```{r message=F, warning=F}
#plot.forecasts(forecast.seasons, seasonNumber, obs.sx, maxInc, fig)
plot.forecasts(forecast.seasons, 1, obs.s1, 300, "fig_S11")
plot.forecasts(forecast.seasons, 2, obs.s2, 300, "fig_S11")
plot.forecasts(forecast.seasons, 3, obs.s3, 300, "fig_S11")
plot.forecasts(forecast.seasons, 4, obs.s4, 300, "fig_S11")
```


Here, we can save our progress.
```{r}
save.image(file="../cache/Table_S1_and_Figures_S9_S11.RData") 
```